---
title: "OLIGO_mofa"
format: html
editor: visual
---

Pseudobulked data is loaded. Script used for this was the "MOFA_OLIGOprep1.R". I think, should verify
```{r}
load("/scratch/prj/bcn_marzi_lab/analysis_cutandtag_pd_sc/student_data_package/jd_analysis_sc/saved_objects/mofa_OLIGO_pseudobulked_granges.RData")
setwd("/scratch/prj/bcn_marzi_lab/analysis_cutandtag_pd_sc/student_data_package/jd_analysis_sc")
```

```{r}
dim(bulk_counts)
dim(opalin_vst)

colnames(bulk_counts)
colnames(opalin_vst)
```

create the input list for opalin subtype and oligo bulk data
```{r}
mofa_input_opalin <- list(
  bulk   = as.matrix(bulk_counts),
  opalin = as.matrix(opalin_vst)
)
```

create the moda object and plot the data overview 
```{r}
library(MOFA2)
mofa_opalin <- create_mofa(mofa_input_opalin)
```
```{r}
# plot data overview
plot_data_overview(mofa_opalin)
```
Get data options

```{r}
data_opts <- get_default_data_options(mofa_opalin)
data_opts
```
```{r}
# model options
model_opts <- get_default_model_options(mofa_opalin)
model_opts$num_factors <- 3
model_opts
```
```{r}
# training options
train_opts <- get_default_training_options(mofa_opalin)
train_opts$convergence_mode <- "slow"
train_opts$seed <- 42


train_opts
```
prepare the MOFA model and train
The model had issues with using less than 10 samples, as recommnded I have reduced my factors down to 2 from 10

```{r}
mofa_opalin <- prepare_mofa(mofa_opalin,
data_options     = data_opts,
model_options = model_opts,
training_options = train_opts)

library(reticulate)
library(basilisk)

mofa_opalin <- run_mofa(
  mofa_opalin,
  outfile = "mofa_opalin_model.hdf5",
  use_basilisk = TRUE
)
```

Add the metdata into what is called a slot. This is where the MOFA object hold relevant data and indo .Ex input data, meta data, expectations 
```{r}
names(mofa_opalin@data)
dim(mofa_opalin@data$bulk$group1)
```


```{r}
# sanity check
stopifnot(all(sort(sample_meta$sample) == sort(unlist(samples_names(mofa_opalin)))))

# add metadata
samples_metadata(mofa_opalin) <- sample_meta
```

```{r}
plot_factor_cor(mofa_opalin)
```




```{r}
plot_variance_explained(mofa_opalin, max_r2 = 15)
```


