---
title: "OLIGO_mofa"
format: html
editor: visual
---

Pseudobulked data is loaded. Script used for this was the "MOFA_OLIGOprep1.R". I think, should verify
```{r}
load("/scratch/prj/bcn_marzi_lab/analysis_cutandtag_pd_sc/student_data_package/jd_analysis_sc/saved_objects/mofa_OLIGO_pseudobulked_granges.RData")
setwd("/scratch/prj/bcn_marzi_lab/analysis_cutandtag_pd_sc/student_data_package/jd_analysis_sc")
```

```{r}
dim(bulk_counts)
dim(opalin_vst)

colnames(bulk_counts)
colnames(opalin_vst)
```

Selecting the top 5000 highly variable features (though this feels low)
```{r}
# select top 5000 most variable peaks per view
bulk_vars       <- apply(bulk_counts, 1, var)
bulk_top_idx    <- head(order(bulk_vars, decreasing = TRUE), 5000)
bulk_hvf        <- bulk_counts[bulk_top_idx, ]
bulk_hvf        <- bulk_hvf - rowMeans(bulk_hvf)

# opalin
opalin_vars     <- apply(opalin_vst, 1, var)
opalin_top_idx  <- head(order(opalin_vars, decreasing = TRUE), 5000)
opalin_hvf      <- opalin_vst[opalin_top_idx, ]
opalin_hvf      <- opalin_hvf - rowMeans(opalin_hvf)

dim(bulk_hvf)
dim(opalin_hvf)

```








create the input list for opalin subtype and oligo bulk data
```{r}
mofa_input_opalin <- list(
  bulk   = as.matrix(bulk_hvf),
  opalin = as.matrix(opalin_hvf)
)

mofa_opalin <- create_mofa(mofa_input_opalin)
plot_data_overview(mofa_opalin)
```

create the moda object and plot the data overview 
```{r}
library(MOFA2)
mofa_opalin <- create_mofa(mofa_input_opalin)
```
```{r}
# plot data overview
plot_data_overview(mofa_opalin)
```
Get data options

```{r}
data_opts <- get_default_data_options(mofa_opalin)
data_opts
```
```{r}
# model options
model_opts <- get_default_model_options(mofa_opalin)
model_opts$num_factors <- 3
model_opts
```
```{r}
# training options
train_opts <- get_default_training_options(mofa_opalin)
train_opts$convergence_mode <- "slow"
train_opts$seed <- 42


train_opts
```
prepare the MOFA model and train
The model had issues with using less than 10 samples, as recommnded I have reduced my factors down to 3 from 10

```{r}
mofa_opalin <- prepare_mofa(mofa_opalin,
data_options     = data_opts,
model_options = model_opts,
training_options = train_opts)

library(reticulate)
library(basilisk)

mofa_opalin <- run_mofa(
  mofa_opalin,
  outfile = "mofa_opalin_model.hdf5",
  use_basilisk = TRUE
)
```

Add the metdata into what is called a slot. This is where the MOFA object hold relevant data and indo .Ex input data, meta data, expectations 
```{r}
names(mofa_opalin@data)
dim(mofa_opalin@data$bulk$group1)
```


```{r}
# sanity check
stopifnot(all(sort(sample_meta$sample) == sort(unlist(samples_names(mofa_opalin)))))

# add metadata
samples_metadata(mofa_opalin) <- sample_meta
```

```{r}
plot_factor_cor(mofa_opalin)

plot_variance_explained(mofa_opalin, max_r2 = 15)
```




```{r}

```
Need to do some checks, looking at an explanation for the issues with Factor 1
```{r}
# check if there are big differences in total accessibility per sample
colSums(bulk_counts)
colSums(opalin_vst)
```



Going to try all the olgio subtypes vs bulk so 4 views and see what happens over 2 factors. As recommended I have reduced the amount of factors and adjusted maxiter to increase the number of iterations. Doubled the deafualt 

```{r}
# select HVFs for bulk and 3 subtypes

bulk_vars       <- apply(bulk_counts, 1, var)
bulk_top_idx    <- head(order(bulk_vars, decreasing = TRUE), 5000)
bulk_hvf        <- bulk_counts[bulk_top_idx, ]
bulk_hvf        <- bulk_hvf - rowMeans(bulk_hvf)


opalin_vars     <- apply(opalin_vst, 1, var)
opalin_top_idx  <- head(order(opalin_vars, decreasing = TRUE), 5000)
opalin_hvf      <- opalin_vst[opalin_top_idx, ]
opalin_hvf      <- opalin_hvf - rowMeans(opalin_hvf)

plekhg1_vars    <- apply(plekhg1_vst, 1, var)
plekhg1_top_idx <- head(order(plekhg1_vars, decreasing = TRUE), 5000)
plekhg1_hvf     <- plekhg1_vst[plekhg1_top_idx, ]
plekhg1_hvf     <- plekhg1_hvf - rowMeans(plekhg1_hvf)

opc_vars        <- apply(opc_vst, 1, var)
opc_top_idx     <- head(order(opc_vars, decreasing = TRUE), 5000)
opc_hvf         <- opc_vst[opc_top_idx, ]
opc_hvf         <- opc_hvf - rowMeans(opc_hvf)

# create 4-view MOFA object
mofa_input_all <- list(
  bulk    = as.matrix(bulk_hvf),
  opalin  = as.matrix(opalin_hvf),
  plekhg1 = as.matrix(plekhg1_hvf),
  opc     = as.matrix(opc_hvf)
)

mofa_all <- create_mofa(mofa_input_all)
plot_data_overview(mofa_all)
```



```{r}
data_opts_all  <- get_default_data_options(mofa_all)

model_opts_all <- get_default_model_options(mofa_all)
model_opts_all$num_factors <- 2

train_opts_all <- get_default_training_options(mofa_all)
train_opts_all$convergence_mode <- "slow"
train_opts_all$maxiter <- 2000
train_opts_all$seed <- 42

mofa_all <- prepare_mofa(mofa_all,
                         data_options     = data_opts_all,
                         model_options    = model_opts_all,
                         training_options = train_opts_all)

mofa_all <- run_mofa(mofa_all,
                     outfile      = "mofa_all_model.hdf5",
                     use_basilisk = TRUE)
```

